Blueprint for nuclear spin characterization in single color center

### 2023-08-03 article.tex
a
fig

\title{Blueprint for nuclear spin characterization in color centers}% Force line breaks with \\
%\thanks{A footnote to the article title}%

@@

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
%\tableofcontents
\section{Introduction and background}
Optically active defects in solids, color centers, has been applied for realizing different quantum applications \cite{awschalom2018quantum}, such as a quantum network \cite{pompili2021realization}, quantum sensing \cite{degen2017quantum}, and quantum register \cite{bradley2019ten}. Each center contains an electron spin that can be controlled directly with microwaves and optics and many nuclear spins that can be manipulated through the electron spin. Characterizing nuclear spins is necessary for enlarging the quantum register and precise description of the electron spin dynamics. Traditionally, nuclear spin characterization is done with Opticall Detected Magnetic Resonance (ODMR) (ref to Phillip Neumann's works), however the spectroscopic resolution is limited to the $T_2^\ast$ of the electron spin. Hence, some sequence is needed to increase the resolution. People have used Hahn Echo sequence to refocus the electron spin and enhance the coherence time of the electron spin \cite{childress2006coherent}. But this method only works for defects with a specific nuclear spins configuration. Depending on the applied protocol for nuclear spins characterization, the number of accessible nuclear spin can increase. Electron Spin Echo Envelop Modulation (ESEEM) types of sequences allow us to access the highest number of nuclear spins since the spectroscopy resolution is limited to the longitudinal relaxation time (T1) of the system (see fig. \ref{fig:1}). 
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{pict/fig1.png}
		\caption{a. nuclear spins that can be characterized using each sequence. b. number of nuclear spins as a function of distance for the defect. The dashed line shows having T1 = 1s limitation, how many nuclear spins are accessible in diamond. }
		\label{fig:1}
	\end{center}
\end{figure*}

@@

\begin{figure*}%[htbp]

@@

We want to make a sequence that is limited by T1 instead of T2. The idea of 5-pulse ESEEM \cite{kasumaj20085} is that to create an entanglement (Hahn-echo) before the free evolution (Ramsey). In other words, change the initial density matrix for the Ramsey sequence from such that entanglement already exist in the electron polarization terms of the density matrix. Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos that are separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. The analytical formula to this sequence is provided in the appendix \ref{app:5pESEEM}. During the free evolution time, each nuclear spins oscillates with either of two resonance frequencies obtained from eq. \ref{eq:H}.
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/fig4.png}
		\caption{a. FFT of ESEEM signal b. a zoom into weakly coupled nuclear spins c. Enhancing weakly coupled nuclear spins by having 72 pulses in each entangling period. $\tau$ is set to the Larmor bright spot.}
		\label{fig:4}
	\end{center}
\end{figure*}
Fig \ref{fig:4} shows the furrier transform of the signal. Zooming into area close to the Larmor frequency reveals the weakly coupled nuclei. Each nuclear spins have two peaks in the spectrum. Expanding the resonance frequencies up to first order with respect to hyperfine coupling results that the weakly coupled nuclei keep their order as they go further away from the nuclei. This fact makes their characterization rather straight forward. To observe the weakly coupled nuclei more pronouncedly, one can use DDESEEM sequence by adding more $\pi$ pulses in the entangling periods. To distinguish other nuclei, extra measurements is required. One method is to use two dimensional hyperfine correlation spectroscopy (2D Hyscore)  \ref{vorobyov2022addressing}. However, this method is rather time-consuming as two time variables has to be swept. We suggest sweeping $\tau$ parameter and keep the track of frequency amplitude as a function of $\tau$. Two frequencies that belong to the same nuclear are correlated because the modulation depth oscillates with blind spot terms of both frequencies $\sin^2(\frac{\omega_0 \tau}{2}) \sin^2(\frac{\omega_1 \tau}{2})$. Fig. \ref{fig:5} shows two frequencies goes to bright and blind spots simultaneously. Hence, taking two dimensional correlation of the spectrum, one can deduce which two frequencies are correlated and belong to the same nuclear. \\

>>>>

\title{Blueprint for nuclear spin characterization in single color center}% Force line breaks with \\
%\thanks{A footnote to the article title}%

@@

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
%\tableofcontents
\section{Introduction and background}
Optically active defects in solids, color centers, has been applied for realizing different quantum applications \cite{awschalom2018quantum}, such as a quantum network \cite{pompili2021realization}, quantum sensing \cite{degen2017quantum}, and quantum register \cite{bradley2019ten}. Each center contains an electron spin that can be controlled directly with microwaves and optics and many nuclear spins that can be manipulated through the electron spin. Characterizing nuclear spins is necessary for enlarging the quantum register and precise description of the electron spin dynamics. Traditionally, nuclear spin characterization is done with Opticall Detected Magnetic Resonance (ODMR) (ref to Phillip Neumann's works), however the spectroscopic resolution is limited to the $T_2^\ast$ of the electron spin. Hence, some sequence is needed to increase the resolution. People have used Hahn Echo sequence to refocus the electron spin and enhance the coherence time of the electron spin \cite{childress2006coherent}. But this method only works for defects with a specific nuclear spins configuration. Depending on the applied protocol for nuclear spins characterization, the number of accessible nuclear spin can increase. Electron Spin Echo Envelop Modulation (ESEEM) types of sequences allow us to access the highest number of nuclear spins since the spectroscopy resolution is limited to the longitudinal relaxation time (T1) of the system (see fig. \ref{fig:1}). 
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{pict/drawing0.pdf}
		\caption{a. nuclear spins that can be characterized using each sequence. b. number of nuclear spins as a function of distance for the defect. The dashed line shows having T1 = 1s limitation, how many nuclear spins are accessible in diamond. }
		\label{fig:1}
	\end{center}
\end{figure*}

@@

\begin{figure*}[htbp]

@@

We want to make a sequence that is limited by T1 instead of T2. The idea of 5-pulse ESEEM \cite{kasumaj20085} is that to create an entanglement (Hahn-echo) before the free evolution (Ramsey). In other words, change the initial density matrix for the Ramsey sequence from such that entanglement already exist in the electron polarization terms of the density matrix. Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos that are separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. The analytical formula to this sequence is provided in the appendix \ref{app:5pESEEM}. During the free evolution time, each nuclear spins oscillates with either of two resonance frequencies obtained from eq. \ref{eq:H}.
\begin{figure*}[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/fig4.png}
		\caption{a. FFT of ESEEM signal b. a zoom into weakly coupled nuclear spins c. Enhancing weakly coupled nuclear spins by having 72 pulses in each entangling period. $\tau$ is set to the Larmor bright spot.}
		\label{fig:4}
	\end{center}
\end{figure*}
Fig \ref{fig:4} shows the furrier transform of the signal. Zooming into area close to the Larmor frequency reveals the weakly coupled nuclei. Each nuclear spins have two peaks in the spectrum. Expanding the resonance frequencies up to first order with respect to hyperfine coupling results that the weakly coupled nuclei keep their order as they go further away from the nuclei. This fact makes their characterization rather straight forward. To observe the weakly coupled nuclei more pronouncedly, one can use DDESEEM sequence by adding more $\pi$ pulses in the entangling periods. To distinguish other nuclei, extra measurements is required. One method is to use two dimensional hyperfine correlation spectroscopy (2D Hyscore)  \cite{vorobyov2022addressing}. However, this method is rather time-consuming as two time variables has to be swept. We suggest sweeping $\tau$ parameter and keep the track of frequency amplitude as a function of $\tau$. Two frequencies that belong to the same nuclear are correlated because the modulation depth oscillates with blind spot terms of both frequencies $\sin^2(\frac{\omega_0 \tau}{2}) \sin^2(\frac{\omega_1 \tau}{2})$. Fig. \ref{fig:5} shows two frequencies goes to bright and blind spots simultaneously. Hence, taking two dimensional correlation of the spectrum, one can deduce which two frequencies are correlated and belong to the same nuclear. \\
